
---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bayescores: Comprehensive quantification of clinical benefit in randomized controlled trials using Bayesian AFT cure models

**`bayescores`** provides a comprehensive toolkit for analyzing randomized controlled trials (RCTs). This package introduces the **Bayesian Clinical Benefit Scores (BayeScores)**, a novel metric to quantify clinical benefit by accounting for both survival prolongation and cure rates.

The package includes functions to:

- Simulate realistic survival data from a mixture cure model.
- Fit Bayesian Accelerated Failure Time (AFT) mixture cure models using Stan.
- Visualize model results and diagnostics.
- Calculate and visualize BayeScores to provide a holistic measure of clinical benefit.

## Installation

Install the development version of `bayescores` from GitHub:

```r
# install.packages("devtools")
devtools::install_github("albertocarm/bayescores")
```

## Full example

This example illustrates the complete workflow—from data simulation to visualization of benefit scores. Here, simulated RCTs enable clear observation of how parameter changes influence BayeScores.

### Step 1: Load the package

```{r load-library}
library(bayescores)
```

### Step 2: Simulate survival data

Simulate a 300-patient trial with long-term survival fractions of 40% (experimental) and 20% (control), with median survival of 12 months in the control arm. The experimental treatment extends survival by 50% among non-cured patients.

```{r simulate-data}
# to work with real data, you must digitize the curves from the original paper using tools such as WebPlotDigitizer (https://automeris.io/)
# and apply the “reconstructing individual-level data” method to invert the Kaplan–Meier equations; see a tutorial at https://pharmasug.org/proceedings/2024/RW/PharmaSUG-2024-RW-125.pdf


set.seed(123)

sim_data <- simulate_weibull_cure_data(
  n_patients = 300,
  cure_fraction_ctrl = 0.20,
  cure_fraction_exp = 0.40,
  max_follow_up = 60,
  weibull_shape = 1.2,
  median_survival_ctrl = 12,
  time_ratio_exp = 1.5
)

plot_km_curves(sim_data)

data(package = "bayescores")
```

### Step 3: Simulate study toxicity

Generate toxicity data for a trial with 1:1 randomization (300 control patients), baseline toxicity of 50% (any-grade), 20% severe events (G3–4), and 1.5× higher toxicity in the experimental arm. QoL outcomes assume "Significant Improvement" with "Very High" evidence.

**Quality of Life Parameters:**

- `qol_scenario` (expected QoL outcome):
  - `1`: **Significant Improvement**
  - `2`: **Stabilization / Probable Benefit**
  - `3`: **No Difference / Marginal Benefit**
  - `4`: **Deterioration**
  - `5`: **Insufficient Data / Unknown**

- `qol_strength` (confidence in QoL evidence):
  - `1`: **Very Low**
  - `2`: **Low**
  - `3`: **Moderate**
  - `4`: **High**
  - `5`: **Very High**

```{r simulate-toxicity}
toxicity_trial <- simulate_trial_data(
  n_control = 300,
  ratio_str = "1:1",
  control_g1_4_pct = 50,
  control_g3_4_pct = 20,
  tox_ratio = 1.5,
  qol_scenario = 1,
  qol_strength = 5
)
```

Visualize toxicity with AMIT plots:

**Any-grade toxicity (Grades 1–4)**

```{r plot-any-grade-toxicity}
create_amit_plot(
  trial_object = toxicity_trial,
  grade_type = "any_grade",
  main_title = "Example: Any-Grade Toxicity Profile",
  data_element = "toxicity",
  n_element = "N_patients"
)
```

**Severe toxicity (Grades 3+)**

```{r plot-severe-toxicity}
create_amit_plot(
  trial_object = toxicity_trial,
  grade_type = "severe_grade",
  main_title = "Example: Severe-Grade (G3+) Toxicity Profile",
  data_element = "toxicity",
  n_element = "N_patients"
)
```

### Step 4: Fit the Bayesian cure model

Fit the Bayesian AFT cure model (use higher `iter` in practice):

```{r fit-model}
bayesian_fit <- fit_bayesian_cure_model(
  sim_data,
  time_col = "time",
  event_col = "event",
  arm_col = "arm",
  iter = 2500,
  chains = 4
)
```

### Step 5: Analyze and visualize model results

The mixture cure model separates individuals into:

- **Cured (long-term survivors)**: negligible event risk long-term.
- **Susceptible (uncured)**: ongoing event risk; survival prolonged but not cured.

Inspect numerical summaries and visualize posterior distributions. You can verify that the model satisfactorily recovers the time ratio and the fractions of long‑term survivors:

```{r numerical-summary}
print(bayesian_fit$stan_fit, pars = c("beta_cure_arm", "beta_surv_arm", "alpha"))
outcomes(bayesian_fit)
```

**Posterior distributions**

```{r plot-parameters, fig.cap="Figure 1: Posterior density distributions for Time Ratio, Odds Ratio of Cure, and Cure Probability Difference."}
plot_densities(bayesian_fit)
```

**Posterior predictive check**
You can observe how the model’s predictions align satisfactorily with the Kaplan–Meier estimator:

```{r plot-model-fit, fig.cap="Figure 2: Posterior predictive check (model vs Kaplan-Meier data)."}
plot(bayesian_fit)
```

### Step 6: Integrate toxicity data

Toxicity is summarized by a burden‑of‑toxicity score, which weights both the severity and the category of toxicity (see technical documentation). Estimates below zero, as observed here, indicate greater toxicity in the experimental arm, consistent with the simulation:

```{r toxicity-integration, fig.cap="Figure 3: Distribution of burden-of-toxicity score."}
toxicity_output <- calculate_toxicity_adjustment(
  trial_data = toxicity_trial,
  n_simulations = bayesian_fit$n_draws
)

plot_toxicity_adjustment(toxicity_output)
```

### Step 7: Quality-of-life weighting

QoL adjustments modeled using multinomial distribution:

```{r qol-weighting, fig.cap="Figure 4: Multinomial distribution of QoL levels."}
qol_scores <- sample_qol_scores(
  prob_vector = toxicity_trial$qol,
  n_samples = bayesian_fit$n_draws
)

plot_qol_histogram(qol_scores)
```

We have chosen this methodology because quality‑of‑life data are published very inconsistently—indeed, the data are a mess—and the approach will likely evolve over time. Therefore, you must read the paper, qualitatively interpret the results, and assess both the impact on quality of life and the strength of evidence. The outcome will be a multinomial distribution, as shown below. It is quite straightforward: to work with real data, the package includes the function generate_qol_vector(), which interactively generates a quality‑of‑life (QoL) probability vector.


### Step 8: Extract posterior samples and compute BayeScores

Okay, we have everything: efficacy, toxicity, QoL. It’s time to extract posterior samples and compute BayeScores.

To obtain utility, a cumulative logistic function is used, anchored by the minimum clinically relevant benefit values, which we—two oncologists—have agreed upon as the time ratio of 1.15 (e.g., extending OS from 10 to 11.5 months) or a long‑term survival increase of 4 %.

Inspired by multi‑attribute utility theory (MAUT), we have also chosen the thresholds at which drugs become not only active but promising: a time ratio of 1.35 and a survival‑rate difference of 12 %.

Now, look at the output. To derive final utilities, an adaptive policy is applied: the function has detected that this clinical trial increases the long‑term survivor rate and has adjusted the weighting scores relative to the predefined ones (see technical specification).


```{r calculate-scores}
efficacy_inputs <- list(
  tr_posterior_samples = extract_mcmc_time_ratios(bayesian_fit),
  cure_posterior_samples = extract_mcmc_cure_diffs(bayesian_fit)
)

final_utilities <- summarize_final_utility(
  efficacy_scores = efficacy_inputs,
  toxicity_scores = toxicity_output$adjustment_vector,
  qol_scores = qol_scores,
  weights_if_no_cure = c(tr=0.7, cure=0.1, tox=0.1, qol=0.1),
  weights_if_cure = c(tr=0.15, cure=0.6, tox=0.15, qol=0.1),
  tr_min_if_no_cure = 1.15,
  tr_hopeful = 1.35,
  cure_min_relevant = 0.04,
  cure_hopeful = 0.12
)

print(final_utilities$component_summary)
```

### Step 9: Visualize final clinical benefit

**BayeScore donut plot**

After these analyses, the tool delivers a final *clinical utility score* derived from all the weightings, representing the drug’s ultimate evaluation. This score is based on a simulation that assumed benefit but also factored in some toxicity, yet still resulted in an improvement in quality of life.

```{r plot-donut, fig.cap="Figure 5: BayeScore donut plot."}
plot_utility_donut(final_utilities)
```

**Final score posterior distribution**

It’s clear that estimating clinical benefit carries substantial uncertainty—studies are noisy, and sample sizes are limited. We’ve rigorously propagated every source of uncertainty throughout the analysis, so that you alone judge the credibility of the results.  

That is the power of the BayeScore: it isn’t a single point estimate but a full Bayesian distribution!  

```{r plot-score-density, fig.cap="Figure 6: Final BayeScore posterior distribution."}
plot_final_utility_density(final_utilities)
```

## Why bayescores? a more meaningful approach

- **Clinically interpretable**: time ratios are more intuitive than hazard ratios; cure models resolve the issue of double counting benefit—first in HR terms and again as the long‑term bonus. this ensures both metrics remain independent.  
- **Accounts for cure scenarios**: explicit modeling of cure rates and long‑term survivor fractions in immunotherapy studies.  
- **Embraces uncertainty**: the Bayesian framework transparently conveys uncertainty.  
- **No thresholds**: there are parameters and weights, but the approach is gradual; it avoids all‑or‑nothing decisions and is transparent and customizable.  

## Citation

```r
citation("bayescores")
```
